---
title: "Test"
author: "Kate Che"
output:
  html_document: 
    toc: yes
    toc_float: yes
  pdf_document: 
    toc: yes
    keep_tex: yes
bibliography: bibl.bib
urlcolor: blue
params:
    filename: glacier.csv
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggvis)
options(scipen = 0, digits = 3)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## Introduction

The data used in study is taken from <https://open.canada.ca/en/open-data>[^1]. The algorithm is taken from @AntonovErmakov_RandomCubaturesQMS.

[^1]: Here is the [direct link](https://open.canada.ca/data/en/dataset/57ba3e50-2024-4570-8acf-4b99544fb689) to data download.

```{r}
glacier <- read.csv("data/glacier.csv", na.strings = "..", comment.char = "#")
names(glacier)[1] <- "REF_DATE"
```

The **data set** *contains* `r nrow(glacier)` measurements of `r nlevels(glacier$GEO)` Canadian ~~glacier~~ mass balance, collected from `r min(glacier$REF_DATE)` till `r max(glacier$REF_DATE)`. Namely, the file includes these glaciers:
```{r}
levels(glacier$GEO)
```

## Hypothesis

We are interested in findong out whether there is a statistically significant cange in mass balance over the observed time period. For these purposes we use **R** (version `r getRversion()`) and an appropriate statistical test called _t-test_: $$t = \frac{\overline{x} - \mu_0}{s/\sqrt{n}}. $$

The workflow is as follow:

1. Read file 
2. Run t-test for each glacier and collect p-values
3. Support the evidence with 
    + a table of results;
    + a plot thet could help demonstrate the effect.
    
## Analysis

```{r analysis, cache=TRUE}
glacier <- glacier %>% 
    select(REF_DATE, GEO, Type.of.measure, VALUE) %>% 
    filter(Type.of.measure == "Annual mass balance") %>% 
    separate(GEO, c("Name", "Location"), sep = ' - ')

# descriptive analysis
g1 <- glacier %>% 
    group_by(Name) %>% 
    summarise(YearsObserved = n(),
              MeanChange = mean(VALUE, na.rm = T),
              WorstChange = min(VALUE, na.rm = T),
              WorstYear = REF_DATE[which.min(VALUE)])

# t-test
g2 <- glacier %>% 
    group_by(Name) %>% 
    do({
        tt <- t.test(.$VALUE, alternative = 'less', mu = 0, conf.level = 0.99)
        data.frame(PValue = tt$p.value, ConfidenceLimit = tt$conf.int[2])
    })

left_join(g1, g2, by = "Name") %>% 
    knitr::kable(caption = "Descriptive statistics and confidence intervals", digits = c(0, 0, 2, 0, 0, 10, 2))
```

```{r ggplot, fig.width=10, fig.height=10}
# ggplot(glacier, aes(REF_DATE, VALUE)) +
#     geom_line() +
#     geom_hline(data = g1, aes(yintercept = MeanChange),
#                color = "red", linetype = "dashed", alpha = 0.8) +
#     facet_wrap(~Name, nrow = 2)
ggvis(glacier, ~REF_DATE, ~VALUE, stroke = ~Name) %>% 
    layer_lines()
```

The conducted analysis supports the evidence that Canadian glaciers are <u>melting</u>.

## Bibliography